---
title: "Homework 8"
author: "Kenna McQueen"
date: "2025-04-14"
output: pdf_document
---

---
title: "Homework 8"
author: "Kenna McQueen"
date: "2025-04-07"
output: pdf_document
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=4, echo=FALSE, fig.width=8, fig.height=5, warning=FALSE, tidy=TRUE, message=FALSE, tidy.opts=list(width.cutoff=60))
```

UT EID: km55667

Github Link: [SDS315](https://github.com/kmcqueen13/SDS315/blob/main/Homework_8.Rmd)

```{r, warning=FALSE}
#Loading in needed libraries
library(tidyverse)
library(mosaic)
library(ggplot2)
library(MatchIt)
```

# **Problem 1: Regression Warm Up**

### **A)**

```{r, results='hide'}
#Reading in the data
creatinine<-read_csv("creatinine.csv")

#linear model of the data
lm_creatinine<-lm(creatclear~age, data=creatinine)
coef(lm_creatinine)
```

The expected creatinine clearance rate expected for a 55-year-old is 113.71 mL/min. The equation $Clearance Rate= 147.81-0.62 \cdot Age$, which was formed by a linear model of clearance rate by age, results in 113.72 when Age=55. 

### **B)**

Creatinine clearance decreases by 0.62 mL/minute per year of age. This was determined from a linear model of creatinine clearance by age. The weight of this model for age is -0.62.

### **C)**

```{r, results='hide'}
#expected for the 40-year-old
147.81-(0.62*40)
#diff in actual-expected for 40 year old
135-123.01

#expected for the 60-year-old
147.81-(0.62*60)
#diff in actual-expected for 60 year old
112-110.61
```

The expected rate given a persons' age is $Clearance Rate= 147.81-0.62 \cdot Age$. Using this, the expected rate of creatinine clearance for a 40-year-old is 123.01. Their actual rate is 135 so the difference, $Actual-Expected$, is $135-123.01=11.99$. The 40-year-old's creatinine clearance rate is 11.99 mL/min more than expected. 

The expected rate of creatine clearance for a 60-year-old is 110.61. The difference between the actual rate of 112 mL/min and the expected rate is $112-110.61=1.39$. Their rate is 1.39 mL/min higher than expected for their age. 

Therefore, the 40-year-old is healthier for their age because their rate is 11.99 mL/min higher than expected, given their age, which is a higher difference than the 60-year-old whose rate is only 1.39mL/min higher than expected for their age.

# **Problem 2: Modeling Disease Growth**

The file covid.csv contains data on daily reported COVID-19 deaths for Italy and Spain—two of the hardest-hit European countries—during the first pandemic wave in February and March of 2020. The columns in this data frame are:
• date: the calendar date
• country: Italy or Spain
• deaths: the number of reported COVID-19 deaths in that country on that day
• days_since_first_death: the number of days elapsed since the first death in that country

Your task is to fit two exponential growth models, one for Italy and one for Spain, using days_since_first_death as the time variable. Use the results of your model to characterize the growth rate and doubling time of the daily death total in each country.

Please include the following in your write-up:
```{r, results="hide"}
#Reading in the dataset
covid<-read_csv("covid.csv")

#An estimated growth rate and doubling time for Italy, with 95% bootstrapped confidence interval for each.
italy_covid<-filter(covid, country=="Italy")

lm_italy<-lm(log(deaths)~days_since_first_death, data=italy_covid)
coef(lm_italy)

boot_italy = do(10000)*lm(log(deaths) ~ days_since_first_death, data=resample(italy_covid))
confint( boot_italy, level = 0.95)

double_italy=mutate(boot_italy, double=70/(days_since_first_death*100))
doubling_italy <- quantile(double_italy$double, c(0.025, 0.975))

#An estimated growth rate and doubling time for Spain, with 95% bootstrapped confidence intervals for each.
spain_covid<-filter(covid, country=="Spain")

lm_spain<-lm(log(deaths)~days_since_first_death, data=spain_covid)
coef(lm_spain)

boot_spain = do(10000)*lm(log(deaths) ~ days_since_first_death, data=resample(spain_covid))
confint(boot_spain, level = 0.95)

double_spain=mutate(boot_spain, double=70/(days_since_first_death*100))
doubling_spain <- quantile(double_spain$double, c(0.025, 0.975))
```

```{r}
#A line graph showing reported daily deaths over time (using days_since_first_death, rather than calendar date, as the relevant time variable) in each country. Your line graph should have two lines, one for each country, distinguished by their color.
ggplot(covid)+geom_line(aes(x=days_since_first_death, y=log(deaths), color=country))+theme_minimal()+labs(color="Country", x="Days Since First Death", y="Log of Death Count", title="Death Counts by Days Since First Death")
```

```{r}
ggplot(covid, aes(x=days_since_first_death, y=log(deaths), color=country)) + 
  geom_smooth(method='lm')+theme_minimal()+labs(color="Country", x="Days Since First Death", y="Log of Death Count", title="Linearized Death Counts by Days Since First Death")
```

The linearized equation for exponential growth is $log(death)=log(\alpha)+\beta \cdot days$, in exponential form it is $death=\alpha \cdot e^{\beta \cdot days}$. 

Using these equations for the death rate in Italy, we get $log(death)=1.019+0.183 \cdot days$ which gives $death= 2.77 \cdot e^{0.183 days}$. Therefore, the estimated growth rate for deaths from Covid in Italy is 0.183 days/log(death). We are 95% confident that the the growth rate is between 0.159 and 0.208. The doubling time for Covid cases is $70/18.3$ which is 3.8 days. A confidence interval at 95% produces an interval from 3.4 to 4.4 days.

Using these equations for the death rate in Spain, we get $log(death)=0.465+0.276 \cdot days$ which gives $death= 1.59 \cdot e^{0.276 days}$. Therefore, the estimated growth rate for deaths from Covid in Spain is 0.276 days/log(death). We are 95% confident that the the growth rate for Covid cases in Spain is between 0.235 and 0.317. The doubling time for this is $70/27.3$ which is 2.5 days. We are 95% confident that the doubling time for Covid cases in Spain is between 2.2 and 3.0 days.

# **Problem 3: Price Elasticity of Demand**

```{r, results='hide'}
milk<-read_csv("milk.csv")

lm_milk = lm(log(sales) ~ log(price), data=milk)
coef(lm_milk)

boot_milk = do(10000)*lm(log(sales) ~ log(price), data=resample(milk))
confint(boot_milk, level = 0.95)
```

The fitted equation on the log scale is $log(Q)=4.72-1.62 \cdot log(P)$ where Q is the sales(quantity demanded) and P is the price. In the power-law model it is $Q=112.2 \cdot P^{-1.26}$. Therefore, the price elasticity of demand for this sample is -1.62. A 95% confident interval for the price elasticity of demand is -1.77 to -1.45. We are 95% confident that when the price of milk increases by 1%, consumers want to buy between 1.77% and 1.45% less of it, on average.
